\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-02 paper/main.tex
Revamp Discussion

To summarize, these results show that the ventricular geometry and the
cross-sectional area of the ventricular ducts clearly impact solute
transport, with notable differences in transport resulting from modest
geometry modifications. These findings may explain the larger
differences in the cilia-deficient computational model results
compared to the experimental data in \Cref{fig:fig4}, \mer{as the
computational geometry is not an exact representation for the
individual zebrafish ventricles.} 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios display $F_i/F_{\mathrm{original}}\times 100\%$.}
    \label{fig:fig6}
\end{figure}

\section{Discussion}
In this study, we have presented a computational model of CSF flow and
solute transport within the brain ventricles.
We used the larval zebrafish brain ventricles since its geometry and
flow mechanisms are well described, thereby allowing us to validate
our simulations experimentally. Nevertheless, our computational framework can be adapted to,
and implemented in, other ventricular systems in the future, once the flow parameters of such systems
are comprehensively resolved. We assumed that intraventricular CSF flow was governed
by the Stokes equations.
By imposing tangential traction and normal pressure
on the CSF flow, we were able to reproduce flow fields with similar features
as observed experimentally. The prominent flow features were vortex-structured flow,
mediated by cilia, and pulsatile flow, mediated by the cardiac cycle.

Using the simulated CSF flow, we modeled solute transport by an advection-diffusion equation.
The transport results compared well with data from in vivo experiments in zebrafish.
Then, by using a computational model as well as zebrafish mutants with paralyzed cilia,
we identified that even though advection by cilia is dominant,
diffusion plays a significant role in protein transport.
Finally, we investigated how local cilia paralysis and ventricular geometries
affect solute distribution. Altogether, we present not only new findings on the role of
diffusion and advection in intraventricular transport, but also provide a novel framework to
study CSF flow and transport in complex geometries.

\subsection{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) over time.
Even though we identified similar dynamics in simulations and
in vivo experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the geometry of the interventricular ducts
impacts solute distribution, and the geometry used for the computational model was larger
than some of the geometries used in experiments.
Notably, there are large variations in ventricular size in the control population at 2 dpf.
Alternatively, dissimilarities in results may relate
to the way we defined the photoconversion site (ROI 1) in the simulations,
since the volume of the photoconversion region determines
the total amount of solute and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight} and
thus under this reported limit, one could assume that Dendra2 may
leak out of the brain ventricles over the time spans we consider
and render the dynamics slower. Finally, an important difference
between the simulations and experiments is the fact that the
computational model is three-dimensional, while the
experiments only resolve two-dimensional features.

Nevertheless, our results convincingly show that
intraventricular transport of proteins depends on a balance of advection and diffusion.
While advection is the strongest transport mechanism on a global scale for
Dendra2 (global Péclet number 13), diffusion is balancing out
advection especially for shorter distances and for regions devoid of motile cilia activity.
Indeed, we identified that the diffusion coefficient had a greater impact on transport to
regions located upstream in the CSF flow field (ROI 3) or in distant ventricles (ROIs 5--6),
than the regions located downstream of cilia-mediated flow (ROI 2).

Meanwhile, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 150 nm radius extracellular vesicles, are almost entirely dependent on advection.
There was essentially zero solute in the posterior ventricle at the
final simulation time (around 860 s), a ventricle where transport is driven
by diffusion due to the absence of motile cilia.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ \textmu m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=2.4$ \textmu m/s over the same
distance would happen on a timescale of $t_a=L/\overline{U}=83$ s.
Comparing this with the minuscule values of $\cbar_6$ after simulating transport
with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

\subsection{Effects of cilia motility and ventricular morphology}
To gain insights into how CSF flow and intraventricular transport depend on cilia
motility and ventricular morphology, we manipulated certain
parameters in our model and the computational geometry.
First, we investigated how complete loss of cilia motility
impacted the transport dynamics. Simulations with a cardiac-induced/no-cilia lead to
slower transport when compared to the baseline model that included cilia. Especially
the middle-ventricular region downstream of CSF flow (ROI 2) and the anterior ventricle (ROI 5)
experienced substantially increased times-to-threshold. We also observed
slower dynamics in the rest of the middle ventricle (ROIs 3 and 4), whereas the
posterior ventricle (ROI 6) was unaffected by the loss of cilia motility,
further indicating that transport from the middle to the posterior ventricle is
purely driven by diffusion at this developmental stage.
We also performed photoconversion experiments on cilia mutant
zebrafish larva (\emph{smh}), observing statistically significant slower transport in 
ROIs 2 and 4. The results provide not only confidence in that our computational model
represents the physics well, but also valuable quantification of the impact of cilia motility
on intraventricular transport. Interestingly, a previous simulation study of reduced cilia motility
also observed a drastic reduction in CSF mixing in specific regions of the brain
ventricles~\cite{Yoshida2022EffectVentricles}, similar to our results. Since impaired brain 
clearance of tracers has been demonstrated in idiopathic normal pressure hydrocephalus
patients~\cite{Eide2020MagneticHydrocephalus}, cilia-enhanced mixing in the brain might be
fundamental to physiological transport within the brain.

Next, we paralyzed cilia in specific regions of
our model. We identified that the cilia in the dorsal middle ventricle contribute
the most to solute distribution, which aligns well with the fact
that these cilia generate the strongest flow field in vivo~\cite{Olstad2019CiliaryDevelopment}
and in our model. In contrast, removing the ventral cilia in the middle ventricle
did not greatly impact flow dynamics, despite these cilia contributing a third of
the total force exerted by the cilia. However, removing ventral cilia in the middle ventricle
lead to reduced transport towards the posterior ventricle,
meanwhile increasing transport towards the anterior ventricle,
as the ventral cilia help induce flow towards the posterior ventricle.
Removing the cilia in the anterior ventricle affected CSF flow only locally,
and did not affect the global transport.

With an interest in assessing the impact of ventricular morphology on solute transport,
we simulated transport in modified ventricular geometries.
Notably, shrinking ducts reduced interventricular solute distribution, while shrinking
the middle ventricle obstructed diffusive transport upstream of the flow field.
We previously identified large interindividual variation in ventricular size and
CSF velocities between larva at given developmental stages and throughout
development~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain, Jeong2024TheZebrafish}.
Our results suggest that these two parameters, CSF velocities and ventricular size,
may well affect each other, which helps explain the variation in the two.

These simulation studies demonstrate that our model is flexible and allows in silico
testing of further hypotheses on the effects of ciliated
lineages and ventricular morphology on CSF flow and intraventricular transport.
We envision that our framework will be useful as a basis for future analyses of
CSF dynamics in more complex geometries and also other animal species and humans.

\subsection{Numerical approximation and computational modeling aspects}
We modeled motile cilia as a stress
exerted tangentially on the CSF at the ventricular walls
while employing a slip boundary condition,
motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles, both previously and in this work, in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the governing equations of the flow
is computationally more expensive
than more conventional approaches such as a
Taylor-Hood discretization
($P_k-P_{k-1}$ elements)~\cite{Stenberg1990ErrorProblem}, or lower-order enriched
elements, such as the mini-element
($P_1-P_1$ with a velocity stabilization)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
This yields $\normlinf{\nabla\cdot\uu_h}=0$ to within solver tolerances,
an important property for the stability of the advection-diffusion
equation\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}.
In addition to its favorable mass-conserving property,
the BDM-DG discretization scheme 
we employed for the Stokes equations
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly without any tailored approach,
since the BDM elements control the normal component of the 
velocity~(\Cref{subsec:appendixA1})~\cite{Brezzi1985TwoProblems}.
If instead (for example) Taylor-Hood
elements were to be used, the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties and increases problem complexity.

The ciliated boundaries in our CSF flow model are regionally homogeneous,
whereas in reality, cilia populate the ventricle walls in a
heterogeneous manner~\cite{Olstad2019CiliaryDevelopment}. Heterogeneous
arrangement of cilia, both in location, beating direction and beating frequency,
has been shown to enhance particle clearance in airway cilia arrays
in mice~\cite{Ramirez-SanJuan2020Multi-scaleArrays}.
Incorporating spatial arrangement heterogeneity in our model could
result in different transport timescales,
but we would not expect notable changes to the general trends of the macroscopic
transport that we study here, since the effects of introducing heterogeneous spacing
of cilia would alter the magnitude of cilia forces, but not their direction.
On the other hand, introducing more heterogeneous beating directionality could
alter the transport timescales as well as the directionality.
In our model, all of the applied cilia
tangential traction is aligned with the rostrocaudal axis, and more heterogeneity
in this alignment to reflect heterogeneity in the beating directionality of
cilia would presumably affect both magnitude and directionality of advective transport.
Previous studies have demonstrated that changes to beating directionality affects transport,
and that specific coordination of the ciliary beating is important in the brain ventricles for
flow and signaling processes~\cite{Eichele2020Cilia-drivenVentricle, Olstad2019CiliaryDevelopment,
Faubel2016Cilia-basedVentricles, Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Pellicciotta2020Cilia},
but also in other phenomena such as maintaining healthy functioning of mucous transport in
airways~\cite{Rayner1996CiliarySyndrome, Schneiter2021Multi-scaleFunction,
Bustamante-Marin2019LackClearance, Tsukita2012CoordinatedFeet} and in nodal flow which establishes
left-right symmetry in mammals~\cite{Yoshiba2014RolesSymmetry, Hirokawa2006NodalAsymmetry,
Sawamoto2006NewBrain}.
Although these studies were mostly on spatial scales resolving individual cilia,
we could envision the same effects in our continuum model,
similar to the results of previous numerical studies~\cite{Ramirez-SanJuan2020Multi-scaleArrays,
Thouvenin2020OriginCanal, Yoshida2022EffectVentricles}.

To simulate protein transport in the brain ventricles, we numerically solved
an advection-diffusion equation. Except for where we imposed a normal pressure
boundary condition on the CSF flow,
we imposed a no-flux condition on the concentration. The normal pressure
boundary condition leads to in- and outflow of CSF,
and this advective transport requires handling of in- and outflux boundary conditions on
the concentration. For the influx, we applied an approximate periodic boundary condition
that conserves the total solute concentration.
We handled the outflux by imposing zero diffusive flux while
leaving the advective flux term as an unknown in the weak form
of the advection-diffusion equation. This approach has been referred to with different names,
such as the ``free boundary condition"~\cite{Papanastasiou1992ACondition},
the ``no boundary condition"~\cite{Griffiths1997TheCondition}, and the 
``consistent flux boundary condition"~\cite{Lynch2020NumericalHemodynamics}. 
In leaving the advective flux term in the weak form we are lacking a boundary condition,
which is problematic in the continuous sense of the problem.
However, the discrete problem is well-posed, because the approach represents
constraining higher-order derivatives of the concentration~\cite{Griffiths1997TheCondition}. 
In defining ROIs 5 and 6 in the anterior and posterior ventricles,
we deliberately excluded the space proximal to
in- and outflow boundaries to limit adverse effects of 
the advective flux boundary condition on the mean concentrations.

A possible improvement to the in- and outflow boundary conditions would be
to consider a 3D-0D coupling of in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
However, if the brain ventricular system in reality is a closed system,
one should rather seek out a way of modeling the CSF flow without open boundaries to avoid
handling in- and outflow boundary conditions. A more appropriate way of handling the beating
cardiac motion could be to impose a ventricular wall deformation, assumed to result from pulsatile
motion of the heart and/or cardiac vessels surrounding the ventricles. This wall deformation would
induce CSF flow while simultaneously permitting enforcement of the impermeability condition
$\uu\cdot\nn=0$ on the ventricular walls. Such wall motion-induced CSF flow has been modeled and
simulated successfully in the human brain ventricular system~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

We heuristically determined the magnitude of the cilia tangential traction boundary condition
with the parameter $\tau = 0.65$ mPa. The maximum stress
$\vert\btau\vert_{\mathrm{max}} = 2.5 \ \mathrm{mPa} = \num{2.5e-3} \ \mathrm{N/m^2}$
was applied in the dorsal-posterior region of the middle ventricle.
Previous simulation studies of cilia-mediated flow in brain ventricles have
used a volumetric force density to represent the cilia.
Based on reconstruction of CSF flow in murine brain ventricles,
a previous study employed a force that increased linearly from
the ventricular wall to a maximum force density of
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$ at a distance 15 \textmu m (the cilium length)
away from the wall~\cite{Siyahhan2014FlowVentricles}.
If we assumed a cilia array of the same thickness
$\delta = 15$ \textmu m in our model, the maximum stress $\vert\btau\vert_{\mathrm{max}}$
would correspond to a maximum force density of 
\begin{equation*}
    f_c = \frac{\vert\btau\vert_{\mathrm{max}}}{\delta}
        = \frac{\num{2.5e-3} \ \mathrm{N/m^2}}{\num{15e-6} \ \mathrm{m}}
        = 169 \ \mathrm{N/m^3},
\end{equation*}
which is smaller than $f_{\mathrm{max}}$, albeit the same order of magnitude.
The length 15 \textmu m is the typical length of cilia in the human brain
\cite{Afzelius2004CiliaRelatedDiseases, Siyahhan2014FlowVentricles},
while in zebrafish they are closer to 5 \textmu m~\cite{Olstad2019CiliaryDevelopment,
Salman2022ComputationalEmbryo, Ringers2023NovelEpithelia}.
With $\delta=5$ \textmu m, our cilia stress corresponds to a maximum force density of
$f_c=507$ N/$\mathrm{m^3}$, which is very similar to
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$~\cite{Siyahhan2014FlowVentricles}.
Another study that used a force density to model ciliary motion is
considered two-dimensional flow in the central canal of zebrafish
embryos~\cite{Thouvenin2020OriginCanal}.
Cilia was modeled using a cilia force density of
4000 N/$\mathrm{m^2}$. Averaging this value over the central canal diameter 8.9 \textmu m
measured in the study yields a maximum force of 449 N/$\mathrm{m^3}$.
It is very interesting to note the similarity 
in cilia force densities estimated in our and these two previous studies.
\fixme{Previous work concluded that motile cilia align by response to the shear
stress of flow. Perhaps there is an optimal cilia stress?}

The diffusion coefficients used in this study are based on experimental values. 
The value $D_3=\num{1.15e-10} \ \mathrm{m^2/s}$
is based on the value $\num{1.15e-10} \pm 0.11 \ \mathrm{m^2/s}$ reported in an
experiment~\cite{GuraSadovsky2017MeasurementExpansion}, while the Stokes-Einstein
relation~(\Cref{eq:D_stokes_einstein})~\cite{Einstein1905UberTeilchen} was used to extrapolate
the value of $D_2$ and estimate the value of $D_1$, based on experimental
data~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells,
Mahmood2023ExosomeTemperature}. The Stokes-Einstein relation assumes the diffusing
substance is of spherical shape, which proteins are not, and the relation does not
necessarily hold in CSF. Moreover, the extrapolation implies the assumption that
the two substances (the one for which we determine $D$ and the one we extrapolate from)
have equal densities. Consequently, there is uncertainty in the diffusion coefficient values used.
We note that one previous study reported that
Dendra2 satisfies the Stokes-Einstein in a solution of similar viscosity
to CSF\cite{GuraSadovsky2017MeasurementExpansion}.
We used CSF density and viscosity values measured in
humans~\cite{Bloomfield1998EffectsFluid},
and because embryonic CSF in zebrafish is high in proteins, nutrients and other substances,
the CSF in which the photoconverted Dendra2 are transported in our experiments may be more viscous
than we assumed. This could reduce the diffusion coefficient (cf.~\Cref{eq:D_stokes_einstein}),
but based on reports of protein concentration having a limited impact on the viscosity of
CSF~\cite{Bloomfield1998EffectsFluid, Brydon1995PhysicalViscosity}, this reduction is presumably
limited. Merely the presence of other proteins and nutrients may also reduce the diffusion
coefficient by reduction of mean-free path lengths of molecules~\cite{Cussler2009Diffusion:Systems}.
Considering the diffusion coefficients $D_1$ and $D_2$ (both smaller
than $D_3$) in our transport analysis provides information about how the simulation
results of Dendra2 transport would be altered if the
diffusion coefficient $D_3$ is overestimated. In the case that $D_3$ is underestimated, 
we expect a trend similar to the difference in the $D_2$ and $D_3$ solute concentrations;
incerasingly diffusive fields.

>>>>

To summarize, these results show that the ventricular geometry and the
cross-sectional area of the ventricular ducts clearly impact solute
transport, with notable differences in transport resulting from modest
geometry modifications. These findings may explain the larger
differences in the cilia-deficient computational model results
compared to the experimental data in \Cref{fig:fig4}, \mer{as the
computational geometry is not an exact representation for the
individual zebrafish ventricles.}

\lyng{We previously identified large interindividual variation in ventricular size and
CSF velocities between larva at given developmental stages and throughout
development~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain, Jeong2024TheZebrafish}.
Our results suggest that these two parameters, CSF velocities and ventricular size,
affect each other.}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios display $F_i/F_{\mathrm{original}}\times 100\%$.}
    \label{fig:fig6}
\end{figure}

\section{Discussion}
In this study, we have presented a computational model of CSF flow and
solute transport within brain ventricles.
We used the larval zebrafish brain ventricles since its geometry and
flow mechanisms are well described, thereby allowing us to validate
our simulations experimentally. Our computational framework can be adapted to
and implemented in other ventricular systems in the future,
once the flow parameters of such systems
are adequately resolved.
We assumed that intraventricular CSF flow was governed
by the Stokes equations.
By imposing tangential traction and normal pressure
on the CSF flow, we were able to reproduce flow fields with similar features
as observed experimentally. The prominent flow features were vortex-structured flow
mediated by cilia, and pulsatile flow mediated by the cardiac cycle.

Using the simulated CSF flow, we modeled solute transport by an advection-diffusion equation.
The transport results compared well with data from in vivo experiments in zebrafish.
Then, by using a computational model as well as zebrafish mutants with paralyzed cilia,
we identified that even though advection by cilia is dominant,
diffusion plays a significant role in protein transport.
Finally, we investigated how local cilia paralysis and ventricular geometries
affect solute distribution. Altogether, we present not only new findings on the role of
diffusion and advection in intraventricular transport, but also provide a
computational framework to study CSF flow and transport in ventricular geometries.

\subsection{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) over time. Our results convincingly show that
intraventricular transport of solute depends on a balance of advection and diffusion.
While advection is the strongest transport mechanism on a global scale for
Dendra2 proteins (global Péclet number 13), diffusion is balancing out
advection for shorter distances and for regions without motile cilia activity.
For instance, for the larger extracellular vesicles, no solute reached the (distant)
posterior ventricle (a distance of $L = 200$ \textmu m) within the simulation time (around 860 s).
Advective transport by the mean velocity $\overline{U}=2.4$ \textmu m/s over the same
distance would happen on a timescale of $t_a=L/\overline{U}=83$ s, indicating that
transport to the posterior ventricle is solely driven by diffusion at the developmental
stage 2 dpf. The estimated timescale of diffusion $t_d=L^2/D_1$ over the 
same distance is $\approx$24500 s. Larger molecules
transported primarily by advection are thus more easily
compartmentalized within the ventricles
as a result of the compartmentalized CSF flow.

Removing motile cilia slowed down transport within
the middle ventricle and towards the anterior ventricle,
but the transport towards the posterior ventricle was unaffected.
Photoconversion experiments on cilia mutant
zebrafish larva (\emph{smh}) aligned with these observations.
This further indicates that
transport from the middle to the posterior ventricle is
purely driven by diffusion at this developmental stage.
The experimental results provide not only confidence
in the predictive ability of the computational model,
but also quantify the impact of cilia motility
on intraventricular transport.
Interestingly, Yoshida~\emph{et al.} observed a drastic reduction
in CSF mixing in specific regions of the brain
ventricles when simulating reduced cilia motility~\cite{Yoshida2022EffectVentricles}.
\lyng{Since impaired brain 
clearance of tracers has been demonstrated in idiopathic normal pressure hydrocephalus
patients~\cite{Eide2020MagneticHydrocephalus},
cilia-enhanced mixing in the brain ventricles might be
fundamental to physiological transport within the brain.}

\subsection{Model validation and choice of parameters}
Even though we identified similar dynamics in simulations and
in vivo experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the geometry of the
ventricular geometry impacts solute distribution,
and the ventricular size varies substantially 
in the control population at 2 dpf.
Additionally, differences may result
from the definition of the photoconversion site (ROI 1)
as the volume of this region determines the total amount of solute
and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have considered as impermeable in our simulations.
The neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight} and
thus under this reported limit, one could assume that Dendra2 may
leak out of the brain ventricles over the relevant time spans
and render the dynamics slower. 
It may also be the case that our model predicts faster
transport because the Dendra2 diffusion
coefficient we used (based on an experimental value
\cite{GuraSadovsky2017MeasurementExpansion}) is too high.
This could be because embryonic CSF has higher viscosity
than the solution considered in the measurements, 
or the presence of proteins, nutrients and other solutes
that hinder diffusive transport
\cite{Bloomfield1998EffectsFluid, Einstein1905UberTeilchen, Cussler2009Diffusion:Systems}.
Finally, an important difference
is the three-dimensional nature of the computational model,
while the experimental images only show 2D slices.

We modeled motile cilia as a tangential
traction at the ventricular walls
while allowing the CSF to slip,
motivated by experimental observations of
flow fields in zebrafish larval brain ventricles in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The ciliated boundaries in our model are regionally homogeneous,
whereas in reality, cilia populate the ventricle walls in a
heterogeneous manner~\cite{Olstad2019CiliaryDevelopment}. Heterogeneity
in cilia location, beating direction and beating frequency
has been shown to enhance particle clearance in airway cilia arrays
in mice~\cite{Ramirez-SanJuan2020Multi-scaleArrays}.
Other studies have demonstrated that changes to beating directionality affects transport,
and that specific coordination of the ciliary beating is important in the brain ventricles for
flow and signaling processes~\cite{Eichele2020Cilia-drivenVentricle, Olstad2019CiliaryDevelopment,
Faubel2016Cilia-basedVentricles, Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Pellicciotta2020Cilia},
for maintaining healthy functioning of mucous transport in
airways~\cite{Rayner1996CiliarySyndrome, Schneiter2021Multi-scaleFunction,
Bustamante-Marin2019LackClearance, Tsukita2012CoordinatedFeet},
and in the nodal flow that establishes
left-right symmetry in mammals~\cite{Yoshiba2014RolesSymmetry, Hirokawa2006NodalAsymmetry,
Sawamoto2006NewBrain}.
Incorporating spatial heterogeneity in our model would
therefore presumably affect results, and could be
interesting to pursue in further work.

The maximum stress
$\vert\btau\vert_{\mathrm{max}} = 2.5 \ \mathrm{mPa} = \num{2.5e-3} \ \mathrm{N/m^2}$
of the cilia boundary condition
was applied in the dorsal-posterior region of the middle ventricle.
In a model of CSF flow in human brain ventricles,
Siyahhan~\emph{et al.} modeled ciliary motion with
a force that increased linearly from
the ventricular wall to a maximum force density of
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$ at a distance 15 \textmu m (the cilium length)
away~\cite{Siyahhan2014FlowVentricles}.
In a cilia array of thickness $\delta = 15$ \textmu m,
the maximum stress $\vert\btau\vert_{\mathrm{max}}$
in our model would correspond to a maximum force density of 
\begin{equation*}
    f_c = \frac{\vert\btau\vert_{\mathrm{max}}}{\delta}
        = \frac{\num{2.5e-3} \ \mathrm{N/m^2}}{\num{15e-6} \ \mathrm{m}}
        = 169 \ \mathrm{N/m^3}.
\end{equation*}
With $\delta=5$ \textmu m (the typical cilium length in zebrafish brain ventricles
\cite{Olstad2019CiliaryDevelopment,
Salman2022ComputationalEmbryo, Ringers2023NovelEpithelia}) 
our cilia stress corresponds to a maximum force density of
$f_c=507$ N/$\mathrm{m^3}$.
For further comparison, Thouvenin~\emph{et al.}
modeled two-dimensional cilia-mediated 
flow in the central canal of zebrafish
embryos using a force density of 4000 N/$\mathrm{m^2}$\cite{Thouvenin2020OriginCanal}.
Averaging this value over the reported central canal diameter 8.9 \textmu m
yields a maximum force of 449 N/$\mathrm{m^3}$.
It is very interesting to note the similarity 
in cilia force densities estimated in our and these two previous studies.
\fixme{Previous work concluded that motile cilia align by response to the shear
stress of flow. Perhaps there is an optimal cilia stress?}

To simulate protein transport in the brain ventricles we solved
an advection-diffusion equation. We handled advective transport
through the anterior/posterior boundaries
(where cardiac pressure pulsations are imposed)
with an approximate periodic boundary condition
that conserves the total solute concentration.
A possible improvement to this would be
to consider a 3D-0D coupling of in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
Alternatively, the beating cardiac motion could be modeled by ventricular wall deformation
to avoid advective boundary fluxes~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

% The Stokes-Einstein relation assumes the diffusing
% substance is of spherical shape and the relation does not
% necessarily hold in CSF. Moreover, the extrapolation implies the assumption that
% the two substances (the one for which we determine $D$ and the one we extrapolate from)
% have equal densities. We note that one previous study reported that
% Dendra2 satisfies the Stokes-Einstein in a solution of similar viscosity
% to CSF\cite{GuraSadovsky2017MeasurementExpansion}.
% We used CSF density and viscosity values measured in
% humans~\cite{Bloomfield1998EffectsFluid},
% and because embryonic CSF in zebrafish is high in proteins, nutrients and other substances,
% the CSF in which the photoconverted Dendra2 are transported in our experiments may be more viscous
% than we assumed. This could reduce the diffusion coefficient (cf.~\Cref{eq:D_stokes_einstein}),
% but previous works have reported a limited impact of protein concentration on the viscosity of
% CSF~\cite{Bloomfield1998EffectsFluid, Brydon1995PhysicalViscosity}.
% Merely the presence of other proteins and nutrients may also reduce the diffusion
% coefficient by reduction of mean-free path lengths of molecules~\cite{Cussler2009Diffusion:Systems}.

\subsection{Numerical approximation}
The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the Stokes equations
is computationally more expensive
than \fixme{conventional} methods such as a
Taylor-Hood discretization
($P_k-P_{k-1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (such as $P_1-P_1$ with a velocity stabilization)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
This yields $\normlinf{\nabla\cdot\uu_h}=0$,
an important property for the stability of the advection-diffusion
equation that we used to simulate solute transport,
especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}.
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly, since the BDM elements control the normal component of the 
velocity~\cite{Brezzi1985TwoProblems}.
For elements where this is not the case (such as Taylor-Hood elements),
the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties and increases problem complexity
(see \Cref{subsec:appendixA1} for more details).

\subsection{Outlook}
We studied the zebrafish larval brain ventricles,
in which the flow generated by ciliary motion
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment}.
In large brains such as the human brain, ciliary beating mostly 
contributes to flow in localized regions proximal to the ventricular walls
due to the microscopic length of the cilia~\cite{Siyahhan2014FlowVentricles}.
The cilia may play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls,
similar to coral reefs,
where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Flow mediated by cilia indeed serves such a purpose in nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry}, and the discovery
of a cilia-driven flow network in rodent brains suggests such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
Our simulation studies demonstrate that our computational model is
flexible and allows in silico testing of such open questions
and further hypotheses on the effects of ciliated
lineages and ventricular morphology on CSF flow and intraventricular transport.

% Small versus large brain: For instance, a large
% volume would require an extensive amount of time for solutes to distribute
% if based solely on diffusion. Besides, the presence of multiple cilia
% instead of a single cilium per cell -- the former can be found in
% mammalian ependymal cells -- will generate stronger velocities.
% Hence advection may be more dominant in those situations.

% Much evidence indicate that blood vessel pulsations, respiration, CSF secretion
% and brain motion induce the pulsatile bulk flow in the central
% regions of the ventricles~\cite{Vinje2019RespiratoryMeasurements,
% Siyahhan2014FlowVentricles, Linninger2005PulsatileBrain, Enzmann1992BrainImaging,
% Kurtcuoglu2007ComputationalSylvius}.

% To date, motile-cilia deficiency has been associated with altered brain function
% in zebrafish, dysfunctional establishment of left-right symmetry,
% and ventricular enlargement in zebrafish, rodents and humans.~\cite{Afzelius2004CiliaRelatedDiseases,
% Youn2018PrimaryDiseases, Ma2024CiliaDisease, Ibanez-Tallon2002LossHydrocephalus}
